---
title: 'Geog4/6300: Lab 4'
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## One and two sample difference tests

**Overview:**
This lab continues work with the CPS dataset you used in Lab 3. You'll be conducting normality tests, statistical tests of difference, and chi square tests to analyze these data.

###Part 1: Testing the normality of the EARNWEEK variable

We're going to look at the variable for average weekly income: EARNWEEK. Load the IPMUS CPS dataset in this repository using read_csv. Filter these data so you have only valid responses, removing any observations with a value of 9999.99. Now let's test the normality of this variable.

**Question 1** _Load the CPS data and filter it to only valid responses._

```{r}
#Code here

```


**Question 2** _To assess the normality of the EARNWEEK variable, you create a random sample of 5,000 records using the `slice_sample` function (look it up in help for more information). That's because the Shapiro-Wilk test can't handle more than 5,000 records. Create that sample using `slice_sample` and then do the following: 1) Perform a Shapiro-Wilk normality test on this variable. 2) Create a Q-Q plot with a line showing the normal distribution. 3) Create a histogram using ggplot or the base R function._

```{r}
#Code here.

```


**Question 3** _Based on the results you saw in question 2, is this variable normally distributed? How do each of those three tests inform your decision?_

{Your response here}

###Part 2: Testing the differences between states###

Based on the results of the previous questions, test whether the incomes of survey respondents in Atlanta (METAREA code 521) and Chicago (1605) metropolitan areas were equal or different.

**Question 4** _Based on the normality of the data, what is the appropriate statistical test for determining if actual household incomes between these cities are significantly different (p=0.05)? Explain why this test is appropriate and give a null and alternative hypothesis for this test_

{Your response here.}

**Question 5** _Run the test you identified in question 4. Be sure to use the FULL sample for both cities (not the smaller sample from Question 2). Also identify the mean and median values for each dataset, which you can use to determine the nature of the difference--which city (if any) has higher incomes._

```{r}
#Code goes here
```

**Question 6** _Interpret the results of the analysis you did in question 4. Do you reject the null hypothesis? Why or why not? What can we say about household incomes in these two metros, focusing on the significance of the test result and the actual difference in means/medians?_

{Your response here.}

###Part 3: Analyzing differences in rates for use of food assistance

While the U.S. Census regularly provides data on use of SNAP/Food stamp benefits, the CPS provides data on other emergency food sources. Here, you will look at use of emergency food from a church, food pantry, or food bank over the last year (FSFDBNK) in New York, New Jersey, and Pennsylvania.

**Question 7** _Reload the original CPS dataset (so that you're not filtering out based on EARNWEEK as you did in question 1). Now filter for records with valid responses to FSFDBNK (<6 on the scale) that are in the states listed above. Call the head of the table with kable when done_

```{r}
#Code goes here.
```

**Question 8** _Create a contingency table of responses to the FSFDBNK variable by state. This should show the number of observations in each category for each state. In addition, calculate the percentage of the population that used food pantries at all in the last year (categories 2-5) for each state._

```{r}
#Code goes here.
```

**Question 9** _Identify if there are statistically significant differences between these states by conducting a chi square test on the contingency table in question 8. What were your null and alternative hypotheses for the chi square test? Do you reject the null hypothesis based on these results? Why or why not?_

```{r}
#Code goes here
```


{Response goes here.}

**Question 10** _Based on the results of your analysis in questions 8 and 9, what conclusions can you make about food bank usage? Refer to the actual differences in rates between states as well as the significance of the chi-square test in your interpretation._

{Response goes here.}

###Challenge question:
Pick one other variable from this CPS dataset and create an appropriate state level estimate. This will probably involve calculating a rate (e.g., % immigrant) or average (e.g., mean year of immigration).  Map out your estimate using the states_albers geopackage file in the data folder for this lab. 

```{r}
#Code goes here
```

Pick two states and test whether they statistically differ with regards to the variable you created. This will mean testing the normality of the variable, running the appropriate difference test, and interpreting the results. In addition to your code, write a short description of your process, explaining what your results show and why you used the test you did.

```{r}
#Code goes here
```

{Explanation goes here.}

**Disclosure of assistance:** _Besides class materials, what other sources of assistance did you use while completing this lab? These can include input from classmates, relevant material identified through web searches (e.g., Stack Overflow), or assistance from ChatGPT or other AI tools. How did these sources support your own learning in completing this lab?_

{Response here.}

**Lab reflection:** _How do you feel about the work you did on this lab? Was it easy, moderate, or hard? What were the biggest things you learned by completing it?_

{Explanation goes here.}